---
title: "simpleRegress"
author: "Sandra Iovino"
date: "Summer 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

for model length~neck, adjected r^2 = 0.75

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

### Length based on age 

```{r}
ageModel <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(ageModel)
summary.lm(ageModel)
```

Adjusted $r^2$ for the age model is 0.51, this less than the adjusted $r^2$ for neck, so this model is not as good as the neck model.

### Length based on chest

```{r}
chestModel <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(chestModel)
summary.lm(chestModel)
```

Adjusted $r^2$ for chest model is 0.78, this is larger than adjusted $r^2$ for age, so this model is better for the age model.  

### Length based on sex 

```{r}
sexModel <- lm(LENGTH~SEX,data=bears)
plot(LENGTH~SEX,data=bears)
summary.lm(sexModel)
```

Adjusted $r^2$ for sex model is -0.08 this less than adjusted $r^2$ for chest, so this model is not as good for the chest model. 


### Length based on headlen

```{r}
headlenModel <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(headlenModel)
summary.lm(headlenModel)
```

Adjusted $r^2$ for headlen is 0.84, this larger than adjusted $r^2$ for sex, so this model is better for sex model. 


### Length based on headwth

```{r}
headwthModel <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(headwthModel)
summary.lm(headwthModel)
```

Adjusted $r^2$ for headwth is 0.53, this less than adjusted $r^2$ for headlen, so this model is not that good for the headlen model.

### Length based on weight

```{r}
weighthModel <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(weighthModel)
summary.lm(weighthModel)
```

Adjusted $r^2$ for weight is 0.74, this less than adjusted $r^2$ for headwth, so this model is not that good for the headwth model.

so adjusted $r^2$ for headlen was the highest model out of all of them, which was 0.84.


## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 

